---
title: "FACS_histogram_extraction"
author: "Maeva"
date: "5/29/2018"
output: pdf_document
params:
  experiment:
    value: '/Users/maeva/Research/FACS/carmen-20180410'
  sample_pico:
    value: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
library(tidyverse)
library(ggcyto)
library(flowCore)
library(flowViz)
library(openCyto)

```


# Script variables and metadata

```{r, echo=FALSE, message=FALSE}
gating_template <- "gating_schema.txt"
gt_dhc <- gatingTemplate(gating_template)

wd <- params$experiment

d <- str_match(wd, "\\d+")

metadata_full <- 
  dir(wd, pattern = "metadata", full.names = TRUE) %>% 
  read_tsv()

metadata <- 
  metadata_full %>% 
  dplyr::filter(sample_num == params$sample_pico)

base_name <- file.path(wd, "analysis", metadata$name)
```

* Experiment: `r params$experiment`
* Reactor: `r metadata$reactor`
* Replicate: `r metadata$rep`
* Sample tube: `r metadata$sample_num`

# Load flowSet, filter it and Log-transform of FSC and SSC

```{r}
fs <- read.flowSet(path = file.path(wd, "raw_data"),
                   pattern = ".fcs")

fs <- fs[metadata_full %>% 
           dplyr::filter(analyse == 1) %>% 
           .$fs_row]

colnames(fs) <- 
  colnames(fs) %>% 
  str_replace("\\*FSC", "wrong") %>% 
  str_replace("\\*525/50 \\[488\\]", "pico")

colnames(fs)

log10trans <- logTransform("log10 transform")

log10transList <- transformList(c("FSC", "SSC"), log10trans)

```

# Sample frames and gating set creation

```{r}

sample_pico <- params$sample_pico


frame_pico <- 
    metadata %>% 
    dplyr::filter(sample_num == sample_pico) %>% 
    .$fs_row

gs <- 
  fs %>% 
  transform(log10transList) %>% 
  GatingSet()

gating(gt_dhc, gs)

plot(gs)
```

# FSC density - noise gate (decided based on all samples of the same run)

```{r, echo=FALSE, message=FALSE}

p1 <- 
  getData(gs[[frame_pico]]) %>% 
  ggplot(aes(FSC)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_gate(getGate(gs, "cells")[[frame_pico]]) +
  theme_classic()

p2 <- 
  getData(gs[[frame_pico]]) %>% 
  ggplot(aes(FSC, SSC)) +
  geom_hex(bins = 170, show.legend = F) +
  geom_gate(getGate(gs, "cells")[[frame_pico]]) +
  scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11, "Spectral")), trans = "sqrt") +
  theme_classic()

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Remove large cells

```{r, echo=FALSE, message=FALSE}

p1 <- 
  getData(gs[[frame_pico]], "cells") %>%
  ggplot(aes(FSC)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_gate(getGate(gs, "smallCells")[[frame_pico]]) +
  theme_classic()

p2 <- 
  getData(gs[[frame_pico]], "cells") %>%
  ggplot(aes(FSC, SSC)) +
  geom_hex(bins = 170, show.legend = F) +
  geom_gate(getGate(gs, "smallCells")[[frame_pico]]) +
  scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11, "Spectral")), trans = "sqrt") +
  theme_classic()

```

# Remove events with off scale pico signal

```{r, echo=FALSE, message=FALSE}
p1 <- 
  getData(gs[[frame_pico]], "smallCells") %>%
  ggplot(aes(pico)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_gate(getGate(gs, "dhcCells")[[frame_pico]]) +
  theme_classic()

p2 <- 
  getData(gs[[frame_pico]], "smallCells") %>%
  ggplot(aes(FSC, pico)) +
  geom_hex(bins = 170, show.legend = F) +
  geom_gate(getGate(gs, "dhcCells")[[frame_pico]]) +
  scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11, "Spectral")), trans = "sqrt") +
  theme_classic()

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Noise level for pico signal

```{r, echo=FALSE, message=FALSE}

p1 <- 
  getData(gs[[frame_pico]], "dhcCells") %>%
  ggplot(aes(pico)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_gate(getGate(gs, "pico")[[frame_pico]]) +
  theme_classic()

p2 <- 
  getData(gs[[frame_pico]], "dhcCells") %>%
  ggplot(aes(FSC, pico)) +
  geom_hex(bins = 170, show.legend = F) +
  geom_gate(getGate(gs, "pico")[[frame_pico]]) +
  scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11, "Spectral")), trans = "sqrt") +
  theme_classic()

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

# Clean density - 95% confidence interval

```{r}
data <- 
  getData(gs[[frame_pico]], "pico") %>% 
  exprs() %>% 
  as_tibble() %>% 
  select(FSC, pico)

ci95 <- 
  gplots::ci2d(data$FSC, data$pico, method = "bkde2D", show = "none", ci.levels = c(0.95))$contours$`0.95` %>% 
  as_data_frame() %>% 
  rename(FSC = x, pico = y)


filtered_data <- 
  data %>% 
  mutate(keep = sp::point.in.polygon(FSC, pico, ci95$FSC, ci95$pico))

filtered_data %>% 
  sample_frac(0.25) %>% 
  ggplot(aes(FSC, pico)) +
  geom_point(aes(color = as.factor(keep)), size = 0.5) +
  geom_point(data = ci95, color = "red")
```

# Extract histogram

Saved at: `r str_c(base_name, "_histo.tsv")`.

```{r}
p <- 
  filtered_data %>% 
  dplyr::filter(keep == T) %>% 
  select(pico) %>% 
  ggplot(aes(pico)) +
  geom_histogram(bins = 256)

p

r <- 
  ggplot_build(p)$data[[1]] %>% 
  select(count, x) %>% 
  rename(pico = x) %>% 
  mutate(n = row_number())
  
r %>% 
  write_tsv(str_c(base_name, "_histo.tsv"))

```



