---
title: "Histogram_model.Rmd"
author: "Maeva"
date: "5/30/2018"
output: pdf_document
params:
  histo_file:
    value: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## CONSTANTS

```{r, echo=FALSE, message=FALSE}
chr_size <- 1.3

doubling_times <- #doubling times in days
  tibble(reactor = c("A", "Z", "M", "R"),
         doubling_time = c(12.1, 38.8, 48.5, 23.1))
```


## FUNCTIONS 

### Get main peak channel

```{r, echo=FALSE, message=FALSE}
getC1 <- function(data_binned) {
  # Given a df of binned DNA data (channel ~ count), returns the C1 channel
  c1_bin <- 
    data_binned %>% 
    top_n(5, count) %>% 
    .$n %>% 
    mean() %>% 
    round()
  
  return(c1_bin)
}
```


### DNA theoretical distribution

#### Theoretical count in a given channel

```{r, echo=FALSE, message=FALSE}
theoretical_count <- function(bin, start_bin, B, C, tau, total_n, time_per_channel) {
  # Given a c1 channel, B, C, tau and total counts, returns the theoretical counts
  # in a specific channel.
  
  if (bin < start_bin | bin > 2 * start_bin) {
    count_theo <- 0
  } else if (bin == start_bin) {
    count_theo <- total_n * pop_percent(0, B, tau)
  } else if (bin == 2 * start_bin) {
    count_theo <- total_n * pop_percent(B + C, tau, tau)
  } else {
    count_theo <- total_n * pop_percent(B + time_per_channel * (bin - start_bin),
                                    B + time_per_channel * (bin + 1 - start_bin),
                                    tau) # might be missing a bin during intergration
  }
  
  return(count_theo)
}
```

#### Percentage of cells in a given time slice

```{r, echo=FALSE, message=FALSE}
pop_percent <- function(t1, t2, t_total) {
  2 * (2 ^ (-t1 / t_total) - 2 ^ (-t2 / t_total))
}
```


### DNA histogram model for replication in current generation (slow growth)

```{r, echo=FALSE, message=FALSE}

DNAhisto_currentGeneration <- 
  function(histo, B, C, tau, std_dev, std_dev_var, chr_size, plot = FALSE, filename = "") {
  # Input is :
  #       - B: duration of B phase
  #       - C: duration of C phase
  #       - tau: doubling time
  #       - std_dev: std of C1 peak
  #       - std_dev_var: increase of std_dev per channel
  #       - chr_size: chromosome size
  # Optional input:
  #       - plot: should the function plot the model (default: FALSE)
  #       - filename: output file name (default: '')
    
  if ( B + C > tau) {
    stop("DNA replication didn't start in current generation.")
  }
  
  c1_peak <- getC1(histo) 
  total_count <- histo$count %>% sum()
  dna_per_bin <- chr_size / c1_peak
  start_bin <- chr_size / dna_per_bin
  time_per_channel <- C / start_bin
  
  
  histo <-
    histo %>%
    mutate(count_theo = map_dbl(n, 
                            ~theoretical_count(., start_bin, B, C, tau, total_count, time_per_channel))
    )
  
  # std_dev with incremental change
  histo <-
    histo %>% 
    mutate(std_dev = std_dev + std_dev_var * row_number(),
           count_model = map2(n, std_dev, 
                          function(x, y) {histo$count_theo * dnorm(histo$n, x, y) %>% 
                                            as_tibble}) %>% 
                     bind_cols() %>% 
                     summarise_all(sum) %>% 
                     gather(key = x, value = y) %>% 
                     .$y)
  
  deviation <- 
    histo %>% 
    mutate(dif = (sqrt(count) - sqrt(count_model))^2) %>% 
    # filter(n > 0.5 * c1_peak) %>% # ignore the debris / cells without full DNA
    transmute(dif = dif) %>% 
    sum() %>% 
    sqrt(.) / 255
  
  if (plot == TRUE) {

    p <- histo %>% 
      ggplot(aes(x = n / c1_peak)) +
      geom_line(aes(y = count), color = "red") +
      geom_line(aes(y = count_model), color = "blue") +
      annotate("text", x = 2.3, y = max(0.9 * histo$count), 
               label = stringr::str_c("Cell cycle (in days)\nB: ", B,
                                      "\nC: ", formatC(C, digits = 2, format = "f"),
                                      "\nD: ", formatC((tau - C - B), digits = 2, format = "f"),
                                      "\ndeviation: ", formatC(deviation, digits = 2, format = "f"))) +
      labs(x = "DNA content")
    
    ggsave(p, filename = filename)
    
  } else {
    return(tribble(~B, ~C, ~Std_dev, ~Std_dev_var, ~Deviation,
                   B,   C,  std_dev,  std_dev_var,  deviation))
  }
  
}
```

## MODEL FITTING

### Loading histogram

file: `r params$histo_file`

```{r, echo=FALSE, message=FALSE}
histo_raw <- 
  params$histo_file %>% 
  read_tsv()


write(params$histo_file, stdout())

curr_reactor <- 
  params$histo_file %>% 
  str_match("\\/(.)_rep") %>% 
  .[2]

tau <- 
  doubling_times %>% 
  dplyr::filter(reactor == curr_reactor) %>% 
  .$doubling_time
```


### Parameter grid search

```{r}
percent_B <- seq(0, 1, 0.1)
percent_C <- seq(0, 1, 0.1)
std_dev_values <- seq(5, 20, 1)
std_dev_var_values <- seq(0.0, 0.15, 0.03)
```

### Fit

```{r, echo=FALSE, message=FALSE}
fits <- tribble(~B, ~C, ~Std_dev, ~Std_dev_var, ~Deviation)

fits_filename <- params$histo_file %>% str_replace("tsv", "fits")

i <- 0

for (b in percent_B) {

  for (c in percent_C) {

    if (b + c <= 1) {
      
      for (std_dev in std_dev_values) {

        for (std_dev_var in std_dev_var_values) {
            
            fits <- 
              DNAhisto_currentGeneration(histo_raw, 
                                         B = b * tau, 
                                         C = c * tau, 
                                         tau = tau, 
                                         std_dev = std_dev, 
                                         std_dev_var = std_dev_var, 
                                         chr_size = chr_size) %>% 
              bind_rows(fits)
            
            i <- i + 1
            
            if (i %% 10 == 0) {
                fits %>% 
                  arrange(Deviation) %>% 
                  write_tsv(fits_filename, 
                            append = TRUE, 
                            col_names = !file.exists(fits_filename))
              
                fits <- tribble(~B, ~C, ~Std_dev, ~Std_dev_var, ~Deviation)
            }
            
        }
      }
    }
  }
}
```

### Best fit


```{r, echo=FALSE, message=FALSE}
# Reload all fits
all_fits <- 
  fits_filename %>% 
  read_tsv(col_names = c("B", "C", "Std_dev", "Std_dev_var", "Deviation"))


# Plot best fit
best_fit <- 
  all_fits %>% 
  top_n(n = -1, wt = Deviation)

best_fit_plotname <-
  params$histo_file %>% 
  str_replace("tsv", "best_fit.png")


DNAhisto_currentGeneration(histo_raw,
                           B = best_fit$B, 
                           C = best_fit$C, 
                           tau = tau, 
                           std_dev = best_fit$Std_dev, 
                           std_dev_var = best_fit$Std_dev_var, 
                           chr_size = chr_size, 
                           plot = TRUE,
                           filename = best_fit_plotname)
```

  * reactor: `r curr_reactor`
  * tau: `r tau`
  * B: `r best_fit$B * tau`
  * C: `r best_fit$C * tau`
  * D: `r tau - best_fit$B * tau - best_fit$C * tau`
  
  